---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



## Installation

This package requires [cmdstanr](https://mc-stan.org/cmdstanr/articles/cmdstanr.html), and so can't go on CRAN (nor Bioc I think). Install `cmdstanr` first, then use it to install `CmdStan`:

```{r eval=FALSE}
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/",
                                       getOption("repos")))

library(cmdstanr)
check_cmdstan_toolchain()
install_cmdstan(cores = 2)
```

Note that C++ compilers are also required (`check_cmdstan_toolchain()` will check for this). 

Then use `remotes::install_github("andrewGhazi/basehitmodel")` to install the remaining dependencies (listed in the DESCRIPTION file) and `basehitmodel` itself.

## Example

Point `model_proteins_separately` to a directory of mapped basehit counts and it will fit the ZINB model per-protein and save the results in the specified output directory. An optional cache directory can be used to save the pre-processing steps so that future model runs with different statistical parameters (e.g. priors) faster. Help documentation on the other arguments can be displayed with `?model_proteins_separately`. 

```{r eval = FALSE}
model_proteins_separately(count_path = "~/basehit/data/BASEHIT_rescreen/",
                          out_dir = "/path/to/out_dir/",
                          cache_dir = "/path/to/outputs/bh_cache/",
                          split_data_dir = "/path/to/outputs/splits/",
                          ixn_prior_width = 0.15,
                          algorithm = "variational",
                          iter_sampling = 2000,
                          iter_warmup = 1000,
                          save_split = TRUE,
                          save_fits = FALSE,
                          save_summaries = TRUE,
                          bead_binding_threshold = 1,
                          save_bead_binders = TRUE,
                          min_n_nz = 3,
                          min_frac_nz = 0.5,
                          verbose = TRUE,
                          seed = 1234)
```

## Software Versions

This package has been tested on R 4.1.0 and the following package versions (though anything from ~2020 forwards *should* work): 

|pkg       |ver   |
|----------|------|
|dplyr     |1.0.7 |
|data.table|1.14.2|
|magrittr  |2.0.2 |
|openxlsx  |4.2.4 |
|tibble    |3.1.6 |
|tidyr     |1.2.0 |
|cmdstanr  |0.5.1 |
|readxl    |1.3.1 |
|furrr     |0.2.3 |
|stringr   |1.4.0 |
|ggplot2   |3.3.5 |
|patchwork |1.1.1 |
|forcats   |0.5.1 |
|purrr     |0.3.4 |
|posterior |1.2.0 |
|rlang     |1.0.0 |
|stats     |4.1.0 |
|utils     |4.1.0 |
|readr     |2.1.2 |
|progressr |0.10.0|

Any OS that can install R and Stan should be sufficient. We have tested the package most thoroughly on CentOS Linux release 7.6.1810 (Core), as well as various recent versions of macOS and Windows 10.

